Fixing Vulnerabilities Hinders Maintainability

### 2019-02-02 paper/ICPC19/icpc19.tex
rw

ISO describes software maintainability as ``the degree of effectiveness and
efficiency with which a software product or system can be modified to improve
it, correct it or adapt it to changes in environment, and in requirements'' on
software quality ISO/IEC 25010. As ISO does not provide any specific guidelines or
formula to calculate maintainability, we resort on Software Improvement
Group (SIG)'s web-based source code analysis service Better Code Hub
(BCH)~\cite{Visser:2016:OREILLY} to compute the maintainability of a given
project. There are other well-known standards and models that have been proposed to increase software safety: 
Common Criteria~\cite{common:2009} which suffered a lot of criticism regarding the
costs associated and poor technical evaluation; the OWASP 
Application Security Verification Criteria~\cite{oswap:2009} which is focused only
on web applications, and a model proposed by Xu et al. (2013)~\cite{6616351} for 
raiting software security which we believe it was one of the first steps taken
by SIG to introduce security on their maintainability model.
BCH compiles the ISO/IEC 25010 into a set of 10 guidelines that are the
base of BCH's evaluation. Two examples are the
\emph{Write Simple Units of Code} based on the McCabe Complexity~\cite{1702388}
and \emph{Keep your Codebase Small} based on the idea that smaller codebases are
easier to maintain and less prone to defects~\cite{Visser:2016:OREILLY}. Compliance to the guidelines is derived from the SIG's
industry benchmark which, according to official sources, consists of over 8
billion lines of code in more than 180 different technologies. SIG analyzes
around 15 million lines of code, per week.

@@

This empirical study exhibits proof that changes applied in the codebase to fix
vulnerabilities affect code maintainability. Especially, for refactorings that
address \emph{Broken Authentication and Session Management} issues, \emph{Memory
Leaks} and \emph{Denial-of-Service} attacks. There are already several tools that suggest security improvements using static analysis, dynamic analysis, and even artificial intelligence. With this study, we intend to highlight the need for tools to predict the impact of patches on maintainability and documentation
to help developers adopt security best practices~\cite{6311252, 7927935, MESQUIDA201519}.

@@

Previous work focused on coupling metrics has evaluated the impact of refactorings and
exhibited proof that quantifying effectively the impact of refactoring on maintainability
may help to choose the appropriate refactoring type~\cite{1167822}. In contrast to
this work, Hegedus et al.~\cite{HEGEDUS2018313} did not select particular metrics to
assess the effect of refactorings. Instead, statistical tests were used to find
the metrics that have the potential to change significantly after refactorings. 
The differences of maintainability between groups of refactored elements and 
non-refactored were analyzed. And, the authors conclude that the source
code elements subjected to refactorings had significantly lower maintainability
than the elements not affected. Studying the evolution of maintaninability issues during the development of
Android apps, Malavolta et al. (2018)~\cite{8530041} discovered that maintainability 
decreases over time. The present work is based on the 10 guidelines proposed
by the Software Improvement Group (SIG) and focuses solely on evaluating
the impact of patching vulnerabilities on software maintainability.

@@

%
% Previous work studied the effect of programming languages in the quality of
% code found~\cite{ray2014large}. It was found that language design has a
% significant yet moderate impact on software quality. The authors have used the
% number of defects as a construct of software quality. Our work analyzes
% software quality in terms of code maintainability to study how it is affected
% by energy efficiency-code changes.

@@

>>>>

ISO describes software maintainability as ``the degree of effectiveness and
efficiency with which a software product or system can be modified to improve
it, correct it or adapt it to changes in environment, and in requirements'' on
software quality ISO/IEC 25010. As ISO does not provide any specific guidelines or
formula to calculate maintainability, we resort on Software Improvement
Group (SIG)'s web-based source code analysis service Better Code Hub
(BCH)~\cite{Visser:2016:OREILLY} to compute the maintainability of a given
project. There are other well-known standards and models that have been proposed to increase software safety: 
Common Criteria~\cite{common:2009} which suffered a lot of criticism regarding the
costs associated and poor technical evaluation; the OWASP 
Application Security Verification Criteria~\cite{oswap:2009} which is focused only
on web applications, and a model proposed by Xu et al. (2013)~\cite{6616351} for 
raiting software security which we believe it was one of the first steps taken
by SIG to introduce security on their maintainability model.

BCH compiles the ISO/IEC 25010 into a set of 10 guidelines that are the
base of BCH's evaluation. Two examples are the
\emph{Write Simple Units of Code} based on the McCabe Complexity~\cite{1702388}
and \emph{Keep your Codebase Small} based on the idea that smaller codebases are
easier to maintain and less prone to defects~\cite{Visser:2016:OREILLY}. Compliance to the guidelines is derived from the SIG's
industry benchmark which, according to official sources, consists of over 8
billion lines of code in more than 180 different technologies. SIG analyzes
around 15 million lines of code, per week.

@@

This empirical study exhibits proof that changes applied in the codebase to fix
vulnerabilities affect code maintainability. Especially, for refactorings that
address \emph{Broken Authentication and Session Management} issues, \emph{Memory
Leaks} and \emph{Denial-of-Service} attacks. There are already several tools that 
suggest security improvements using static analysis, dynamic analysis, and even 
artificial intelligence. With this study, we intend to highlight the need for tools 
to predict the impact of patches on maintainability~\cite{4724577} and documentation
to help developers adopt security best practices~\cite{6311252, 7927935, MESQUIDA201519}.

@@

\subsection{Software Maintainability}

Previous work focused on coupling metrics has evaluated the impact of refactorings and
exhibited proof that quantifying effectively the impact of refactoring on maintainability
may help to choose the appropriate refactoring type~\cite{1167822}. In contrast to
this work, Hegedus et al.~\cite{HEGEDUS2018313} did not select particular metrics to
assess the effect of refactorings. Instead, statistical tests were used to find
the metrics that have the potential to change significantly after refactorings. 
The differences of maintainability between groups of refactored elements and 
non-refactored were analyzed. And, the authors conclude that the source
code elements subjected to refactorings had significantly lower maintainability
than the elements not affected. Studying the evolution of maintaninability issues during the development of
Android apps, Malavolta et al. (2018)~\cite{8530041} discovered that maintainability 
decreases over time. Palomba et al. (2018)\cite{Palomba:2018:DIM:3231288.3231337} exhibits
proof that code smells should be carefully monitored by programmers, since there is
a high correlation between maintainability aspects and proneness to changes and faults.
The present work is based on the 10 guidelines proposed
by the Software Improvement Group (SIG) and focuses solely on evaluating
the impact of patching vulnerabilities on software maintainability.

@@

Other studies investigated the effect of programming languages in software quality~\cite{Ray:2017:LSP:3144574.3126905,
2019arXiv190110220B}. The first study shows that some programming languages are more buggy than others, i.e., 
language design has impact on software quality. However, the authors of the second
study were not able to reproduce the first one and did not obtain any evidence about the 
impact of language design. Both have used the number of defects as an indicator of
software quality. The present work studies how security refactorings affect software quality
based on the code maintainability analysis.

@@

\pagebreak
